### 2024-03-29 outputs/paper/paper.qmd
model and paper updates

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(arrow)
library(modelsummary)
library(bayesplot)
library(broom)
library(gridExtra)
library(rstanarm)
library(knitr)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-sample
#| fig-cap: Bills of penguins
#| echo: false

analysis_data <- read_parquet(here::here("outputs/data/analysis_data.parquet"))
kable(head(analysis_data))
```

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

distribution_price <-
  ggplot(analysis_data, aes(x = Price)) +
  geom_histogram() +
  theme_minimal()

distribution_EPS <-
  ggplot(analysis_data, aes(x = EPS)) +
  geom_histogram() +
  theme_minimal()

distribution_Dividends <-
  ggplot(analysis_data, aes(x = Dividends)) +
  geom_histogram() +
  theme_minimal()

distribution_Net_Income <-
  ggplot(analysis_data, aes(x = Net_Income)) +
  geom_histogram() +
  theme_minimal()

grid.arrange(distribution_price, distribution_EPS, distribution_Net_Income, distribution_Dividends, nrow = 2)
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| include:false

summary(analysis_data$Price)
summary(analysis_data$EPS)
summary(analysis_data$Net_Income)
summary(analysis_data$Dividends)
```

```{r}
#| label: tbl
#| fig-cap: Relationship between me and ur mom
#| echo: false
#| warning: false
#| message: false

base_net_income_plot <-
  analysis_data |> 
  ggplot(aes(x = Net_Income, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "Net Income",
       y = "Stock Price")

base_dvidends_plot <-
  analysis_data |> 
  ggplot(aes(x = Dividends, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "Dividends",
       y = "Stock Price")

base_eps_plot <-
  analysis_data |> 
  ggplot(aes(x = EPS, y = Price)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "EPS",
       y = "Stock Price")

best_fit_net_income <-
  base_net_income_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )

best_fit_dividends <-
  base_dvidends_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )

best_fit_eps <-
  base_eps_plot +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  )

grid.arrange(best_fit_net_income, best_fit_dividends, best_fit_eps, ncol = 2, nrow = 2)

```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

Simple Model
\begin{align} 
y &= \beta_0 + \beta_1X + \epsilon \\
Y &\sim \mbox{Normal}(\beta, \sigma^2)
\end{align}
Where Y is the stock price and X is earnings per share.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

Multivariable Model
\begin{align} 
y &= \beta_0 + \beta_1X_1 + \beta_2X_2 - \beta_3X_3 + \epsilon \\
Y &\sim \mbox{Normal}(\beta, \sigma^2)
\end{align}
Where X1 is EPS, X2 is net income, and X3 is dividends.

### Model justification
We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

first_model <-
  lm(
    Price ~ EPS,
    data = analysis_data
  )

modelsummary(list("Net_Income" = first_model))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-poopy but hole
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

analysis_data <-
  augment(
    first_model,
    data = analysis_data
  )

distribution_residuals <-
  ggplot(analysis_data, aes(x = .resid)) +
  geom_histogram(binwidth = 1) +
  theme_classic() +
  labs(y = "Number of occurrences", x = "Residuals")

residuals_eps <-
  ggplot(analysis_data, aes(x = EPS, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
  theme_classic() +
  labs(y = "Residuals", x = "Earnings Per Share")

residuals_price <-
  ggplot(analysis_data, aes(x = Price, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
  theme_classic() +
  labs(y = "Residuals", x = "Stock Price")

estimated_vs_actual <-
  ggplot(analysis_data, aes(x = Price, y = .fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme_classic() +
  labs(y = "Estimated Stock Price", x = "Actual Stock Price")

grid.arrange(distribution_residuals, residuals_eps, residuals_price, estimated_vs_actual, nrow = 2)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

  ggplot(analysis_data,aes(y=Price,x=EPS,color=Net_Income)) +
  geom_point() +
  stat_smooth(method="lm",se=FALSE)+
  theme_classic()

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-uuhhhh
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

multi_model <-
  lm(
    Price ~ EPS + Net_Income,
    data = analysis_data
  )

modelsummary(
  list(
    "EPS Only" = first_model,
    "With Net Income" = multi_model
  ),
  fmt = 2
)
```

```{r}
#| include: false
#| eval: true
#| warning: false
model_rstanarm_1 <-
  stan_glm(
    formula = Price ~ EPS + Net_Income,
    data = analysis_data,
    family = gaussian(),
    prior = normal(location = 0, scale = 2.5),
    prior_intercept = normal(location = 0, scale = 2.5),
    prior_aux = exponential(rate = 1),
    seed = 345
  )

saveRDS(
  model_rstanarm_1,
  file = "model_rstanarm_1.rds"
)
```

```{r}
#| include: false
#| eval: true
#| warning: false

model_rstanarm_1 <-
  readRDS(file = "model_rstanarm.rds")
prior_summary(model_rstanarm_1)
```

```{r}
#| include: false
#| eval: true
#| warning: false
model_rstanarm_2 <-
  stan_glm(
    formula = Price ~ EPS + Net_Income,
    data = analysis_data,
    family = gaussian(),
    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
    prior_intercept = normal(0, 2.5, autoscale = TRUE),
    prior_aux = exponential(rate = 1, autoscale = TRUE),
    seed = 1234
  )

saveRDS(
  model_rstanarm_2,
  file = "model_rstanarm_2.rds"
)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-poop
#| tbl-cap: "priors"
#| warning: false

modelsummary(
  list(
    "Non-scaled priors" = model_rstanarm_1,
    "Auto-scaling priors" = model_rstanarm_2
  ),
  fmt = 2
)
```

```{r}
prior_summary(model_rstanarm_2)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(model_rstanarm_2) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(model_rstanarm_2) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
plot(model_rstanarm_2, "trace")

plot(model_rstanarm_2, "rhat")
```

\newpage

# References

### 2024-03-29 scripts/02-data_cleaning.R

@@ -16,6 +16,13 @@ library(arrow)
 #Read in data and save as a data frame
 raw_data <- read_parquet("inputs/data/raw_data.parquet")
 raw_data <- as.data.frame(raw_data)
+largest_companies <- c("MSFT", "AAPL", "NVIDIA", "GOOGL", "AMZN", "META",
+                       "BRK.B", "LLY", "AVGO", "V", "TSLA", "JPM", "WMT",
+                       "UNH", "XOM", "MA", "PG", "HD", "JNJ", "ORCL", "COST",
+                       "MRK", "ABBV", "CRM", "BAC", "CVX", "AMD", "NFLX", "KO",
+                       "PEP", "ADBE", "DIS", "TMO", "MCD", "WFC", "CSCO", "ABT",
+                       "TMUS", "GE", "QCOM", "DHR", "INTU", "CAT", "INTC",
+                       "IBM", "VZ", "AMAT", "CMCSA", "UBER", "AXP")
 
 #Rename variables so they are easier to understand, select the ones needed, and
 #round to two decimal places
@@ -23,9 +30,16 @@ cleaned_data <-
   raw_data |>
   rename("Year" = fyear, "Tic" = tic, "Dividends" = dvc, "EPS" = epsfi, 
          "Net_Income" = ni, "Price" = prcc_f)
+
 cleaned_data <-
   cleaned_data |>
-  select(Year, Tic, Dividends, EPS, Net_Income, Price)
+  select(Year, Tic, EPS, Dividends, Net_Income, Price)
+
+cleaned_data <-
+  cleaned_data |> filter(Dividends >= 0)
+cleaned_data <-
+  cleaned_data |> filter(Tic %in% largest_companies)
+
 cleaned_data$Dividends = round(cleaned_data$Dividends, 2)
 cleaned_data$EPS = round(cleaned_data$EPS, 2)
 cleaned_data$Net_Income = round(cleaned_data$Net_Income, 2)
@@ -37,3 +51,5 @@ head(cleaned_data)
 
 #### Save data ####
 write_parquet(cleaned_data, "outputs/data/analysis_data.parquet")
+write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+

### 2024-03-29 scripts/03-test_data.R

@@ -92,3 +92,4 @@ if (all(analysis_data$EPS |>
 } else {
   "Fail: Not all of the EPS have been cleaned completely"
 }
+

### 2024-03-29 outputs/paper/paper.qmd
updates to the data section

@@
 bibliography: references.bib
+toc: true
 ---
@@
 library(tidyverse)
-library(palmerpenguins)
 library(arrow)
@@
 library(knitr)
+#
+library(gt)
 ```
@@
 
-# Data {#sec-data}
+# Data
+Public companies whose stock are traded on exchanges, must disclose certain financial metrics in their quarterly and yearly financial statements.
+Investors then use these metrics to help inform their decision when selecting companies to invest in.
+Some of the most important metrics investors look at are Earnings Per Share, Net Income, and Dividends [@CiteShakespeare].
+Investments take the form of buying shares or stocks which allows the investor to own a fraction of the company.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The data used for analysis is obtained through Walton Research Data Services (WRDS) by the University of Pennsylvania.
+WRDS provides access to Compustat, a database of financial, statistical, and market information on global companies since 1962.
+Our data comes from Compustat's North America Fundamentals Annual database, which contains both financial data from all public North American companies, collected by Compustat from each company's annual financial statements or from stock exchanges (@fig-Analysis-Data-Sample).
+This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, and @CiteKnitr.
 
 ```{r}
-#| label: fig-sample
-#| fig-cap: Bills of penguins
+#| label: fig-Analysis-Data-Sample
+#| fig-cap: Data from Compustat North America Fundamentals Annual.
+EPS, Dividends, and Net Income are financial statement data while Price are stock exchange data, and Year and Tic can be found on both.
 #| echo: false
@@
 
+## Financial Statement Data
+Earnings per share, net income, and dividends are items that are found on financial statements.
+Specifically, EPS and net income can be found on the income statement while dividends can be found on the statement of retained earnings.
+This data is likely unbiased and free from error as financial statements are required by law to be audited (verified by an independent third party).
+
+EPS is a commonly used measure of a given company's value in USD; it is calculated as net income divided by the number of shares of stock.
+A higher EPS indicates greater value as investors are willing to pay more for a company's shares if they think the company has higher profits relative to its share price [@CiteShakespeare].
+We expect earnings to be positively correlated with stock price.
+
+Net income is used to measure profitability in USD or how much income the company gets to keep after expenses are paid for.
+Higher net income means that the company is profitable by either earning more revenues or reducing expenses.
+Net income is expected to be positively correlated with price as investors want to invest in companies that are more profitable.
+
+## Stock Exchange Data
+The price is the amount that each share costs to buy, it is determined by supply and demand for a company.
+For example, if investors believe a company is profitable, they will buy shares with the goal of eventually sharing in said profit, increasing demand and driving up share price.
+
 ```{r}
-#| label: fig-bills
+#| label: fig-distributions
 #| fig-cap: Bills of penguins
 #| echo: false
+#| warning: false
+#| message: false
 
@@
 
-grid.arrange(distribution_price, distribution_EPS, distribution_Net_Income, distribution_Dividends, nrow = 2)
+grid.arrange(distribution_EPS, distribution_Net_Income, distribution_Dividends, ncol = 3)
 ```
@@
 ```{r}
-#| include:false
+#| echo: false
+#| eval: true
+#| label: tbl-summary-statistics
+#| tbl-cap: " "
+#| warning: false
 
-summary(analysis_data$Price)
-summary(analysis_data$EPS)
-summary(analysis_data$Net_Income)
-summary(analysis_data$Dividends)
+earnings_per_share <- unclass(summary(analysis_data$EPS, digits = 3))
+net_income <- unclass(summary(analysis_data$Net_Income, digits = 7))
+dividends <- unclass(summary(analysis_data$Dividends, digits = 6))
+price <- unclass(summary(analysis_data$Price, digits = 5))
+summary <- data.frame(earnings_per_share, net_income, dividends, price, check.names = FALSE, stringsAsFactors = FALSE)
+
+summary_stats <- c("Min", "1st Quartile", "Median", "Mean", "3rd Quartile", "Max")
+summary <-
+  cols_add(gt(summary), summary_stats) |>
+  cols_move_to_start(summary_stats)
+
+tab_footnote(summary, "Note: ")
 ```

### 2024-03-30 scripts/02-data_cleaning.R
paper updates

@@ -24,8 +24,7 @@ largest_companies <- c("MSFT", "AAPL", "NVIDIA", "GOOGL", "AMZN", "META",
                        "TMUS", "GE", "QCOM", "DHR", "INTU", "CAT", "INTC",
                        "IBM", "VZ", "AMAT", "CMCSA", "UBER", "AXP")
 
-#Rename variables so they are easier to understand, select the ones needed, and
-#round to two decimal places
+#Rename variables so they are easier to understand, select the ones needed
 cleaned_data <-
   raw_data |>
   rename("Year" = fyear, "Tic" = tic, "Dividends" = dvc, "EPS" = epsfi, 
@@ -40,6 +39,12 @@ cleaned_data <-
 cleaned_data <-
   cleaned_data |> filter(Tic %in% largest_companies)
 
+#Create binary variable for dividends
+cleaned_data$Paid_Dividend <- case_when(
+  cleaned_data$Dividends > 0 ~ "Yes", 
+  cleaned_data$Dividends == 0 ~ "No")
+
+#round to two decimal places
 cleaned_data$Dividends = round(cleaned_data$Dividends, 2)
 cleaned_data$EPS = round(cleaned_data$EPS, 2)
 cleaned_data$Net_Income = round(cleaned_data$Net_Income, 2)

### 2024-04-01 outputs/paper/paper.qmd
updates to results section

@@
 title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
+subtitle: "Analysis of the Relationship Between Stock Price and EPS & Dividends of the Largest 50 North American Companies over 10 Years"
+author: Diana Liu
 thanks: "Code and data are available at: LINK."
@@
 bibliography: references.bib
-toc: true
 ---
@@
 library(gt)
+library(grDevices)
 ```
@@
 Our data comes from Compustat's North America Fundamentals Annual database, which contains both financial data from all public North American companies, collected by Compustat from each company's annual financial statements or from stock exchanges (@fig-Analysis-Data-Sample).
-This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, and @CiteKnitr.
+From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years.
+This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr
+
+Similar data sets exist, but WRDS was selected because it hosts the most robust collection in terms of variables and its infrastructure allows users to easily query specific items.
 
@@
 #| label: fig-Analysis-Data-Sample
-#| fig-cap: Data from Compustat North America Fundamentals Annual.
-EPS, Dividends, and Net Income are financial statement data while Price are stock exchange data, and Year and Tic can be found on both.
+#| fig-cap: "Data from Compustat North America Fundamentals Annual.
+EPS, Dividends, and Net Income are financial statement data while Price are stock exchange data, and Year and Tic can be found on both."
 #| echo: false
@@
 ## Financial Statement Data
-Earnings per share, net income, and dividends are items that are found on financial statements.
+Net income, EPS, and dividends are items that are found on financial statements.
 Specifically, EPS and net income can be found on the income statement while dividends can be found on the statement of retained earnings.
@@
 
-EPS is a commonly used measure of a given company's value in USD; it is calculated as net income divided by the number of shares of stock.
-A higher EPS indicates greater value as investors are willing to pay more for a company's shares if they think the company has higher profits relative to its share price [@CiteShakespeare].
-We expect earnings to be positively correlated with stock price.
-
-Net income is used to measure profitability in USD or how much income the company gets to keep after expenses are paid for.
+Net income is used to measure profitability or how much income the company keeps after expenses are paid.
 Higher net income means that the company is profitable by either earning more revenues or reducing expenses.
 Net income is expected to be positively correlated with price as investors want to invest in companies that are more profitable.
+Net income can be negative when the company has lost money.
+In our data set, it is measured in millions USD (x10^6).
+Net income is normally distributed with median of 7120M USD and right skew, mean of 11752 (@fig-distributions).
+Apple (AAPL) had the highest net income in the data set of 99803 achieved in 2022, and Berkshire Hathaway (BRK.B) had the lowest of -22819 in the same year (@tbl-summary-statistics).
+
+```{r}
+#| label: fig-EPS-boxplot
+#| fig-cap: "Boxplots of EPS grouped by dividends"
+#| echo: false
+EPS_boxplot <-
+  analysis_data |> 
+  ggplot(aes(x = Paid_Dividend, y = EPS)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Dividends Paid",
+       y = "EPS")
+
+ylim1 = boxplot.stats(analysis_data$EPS)$stats[c(1, 5)]
+EPS_boxplot = EPS_boxplot + coord_cartesian(ylim = ylim1*1.05)
+EPS_boxplot
+
+```
+
+EPS is a commonly used measure of a given company's value in USD; it is calculated net income divided by the number of shares.
+It is normally distributed and right skewed, with median 4.56 USD and mean 6.2 (@fig-distributions).
+Google (GOOGL) had the highest EPS of 112.2 in 2021, and Tesla (TSLA) had the lowest of -11.83 in 2017, meaning for each share Tesla lost 11.83 USD (@tbl-summary-statistics).
+
+Dividends are distributions of a company's income to the owners of its stock in USD.
+Some companies choose not to pay dividends if they can use their income to reinvest into valuable projects.
+For our analysis we created a binary variable for whether or not a company paid dividends, with any amount greater than zero dollars being Yes.
+Approximately 78% of companies in our data set paid dividends, the largest of which was 25999 USD paid by Caterpillar (CAT) in 2023.
+
+EPS of companies that did not pay dividends has a larger interquartile range than EPS of companies that paid dividends (@fig-EPS-boxplot).
+This means that the observations for companies that paid dividends are more clustered together, likely resulting in smaller standard deviation.
+The mean and median EPS of companies that paid dividends is also higher, suggesting that paying dividends is correlated with having higher EPS (@fig-EPS-boxplot), this makes sense as companies that are profitable enough to have high EPS also have more income to pay out dividends.
+
+```{r}
+#| include: false
+
+count(analysis_data)
+analysis_data %>% filter(Paid_Dividend == "Yes") %>% nrow()
+
+analysis_data %>% filter(Paid_Dividend == "Yes") %>% nrow()/count(analysis_data)
+```
 
@@
 For example, if investors believe a company is profitable, they will buy shares with the goal of eventually sharing in said profit, increasing demand and driving up share price.
+Share price is normally distributed with median of 113.91 USD and mean of 6.2.
+The most expensive stock was Amazon (AMZN) which was worth 3334.34 USD in 2021, and the cheapest stock was Advanced Micro Devices, Inc.
+(AMD) which was worth 2.67 in 2014.
+
+Tic refers to the ticker symbol of each stock.
+It is a unique string of one to five characters used to denote a particular company's stock so that investors know whose stock are being traded on international exchanges across different languages.
+Some tick symbols are similar to the name of the company like AAPL for Apple and GOOGL for Google, but this is not always the case, for example American Express is AXP and Coca-cola is KO.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-summary-statistics
+#| tbl-cap: "Summary statistics of Price, EPS, Net Income, and Dividends of the 50 largest North American Companies by market cap"
+#| warning: false
+
+earnings_per_share <- unclass(summary(analysis_data$EPS, digits = 3))
+net_income <- unclass(summary(analysis_data$Net_Income, digits = 7))
+dividends <- unclass(summary(analysis_data$Dividends, digits = 6))
+price <- unclass(summary(analysis_data$Price, digits = 5))
+summary <- data.frame(earnings_per_share, net_income, dividends, price, check.names = FALSE, stringsAsFactors = FALSE)
+
+summary_stats <- c("Min", "1st Quartile", "Median", "Mean", "3rd Quartile", "Max")
+summary <-
+  cols_add(gt(summary), summary_stats) |>
+  cols_move_to_start(summary_stats)
+
+summary
+```
 
@@
 #| label: fig-distributions
-#| fig-cap: Bills of penguins
+#| fig-cap: "Price, EPS, Net_Income, and Dividends are all normally distributed with right skew"
 #| echo: false
@@
 
-grid.arrange(distribution_EPS, distribution_Net_Income, distribution_Dividends, ncol = 3)
+grid.arrange(distribution_price, distribution_EPS, distribution_Net_Income, distribution_Dividends, ncol = 2)
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-summary-statistics
-#| tbl-cap: " "
-#| warning: false
-
-earnings_per_share <- unclass(summary(analysis_data$EPS, digits = 3))
-net_income <- unclass(summary(analysis_data$Net_Income, digits = 7))
-dividends <- unclass(summary(analysis_data$Dividends, digits = 6))
-price <- unclass(summary(analysis_data$Price, digits = 5))
-summary <- data.frame(earnings_per_share, net_income, dividends, price, check.names = FALSE, stringsAsFactors = FALSE)
-
-summary_stats <- c("Min", "1st Quartile", "Median", "Mean", "3rd Quartile", "Max")
-summary <-
-  cols_add(gt(summary), summary_stats) |>
-  cols_move_to_start(summary_stats)
-
-tab_footnote(summary, "Note: ")
-```
-
-```{r}
-#| label: tbl
-#| fig-cap: Relationship between me and ur mom
-#| echo: false
-#| warning: false
-#| message: false
-
-base_net_income_plot <-
-  analysis_data |> 
-  ggplot(aes(x = Net_Income, y = Price)) +
-  geom_point(alpha = 0.5) +
-  theme_minimal() +
-  labs(x = "Net Income",
-       y = "Stock Price")
-
-base_dvidends_plot <-
-  analysis_data |> 
-  ggplot(aes(x = Dividends, y = Price)) +
-  geom_point(alpha = 0.5) +
-  theme_minimal() +
-  labs(x = "Dividends",
-       y = "Stock Price")
-
-base_eps_plot <-
-  analysis_data |> 
-  ggplot(aes(x = EPS, y = Price)) +
-  geom_point(alpha = 0.5) +
-  theme_minimal() +
-  labs(x = "EPS",
-       y = "Stock Price")
-
-best_fit_net_income <-
-  base_net_income_plot +
-  geom_smooth(
-    method = "lm",
-    se = TRUE,
-    color = "black",
-    linetype = "dashed",
-    formula = "y ~ x"
-  )
-
-best_fit_dividends <-
-  base_dvidends_plot +
-  geom_smooth(
-    method = "lm",
-    se = TRUE,
-    color = "black",
-    linetype = "dashed",
-    formula = "y ~ x"
-  )
-
-best_fit_eps <-
-  base_eps_plot +
-  geom_smooth(
-    method = "lm",
-    se = TRUE,
-    color = "black",
-    linetype = "dashed",
-    formula = "y ~ x"
-  )
-
-grid.arrange(best_fit_net_income, best_fit_dividends, best_fit_eps, ncol = 2, nrow = 2)
-
-```
-
-Talk way more about it.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of our model is to determine the relationship between EPS and stock price first with a simple linear model.
+Then we add a binary variable of whether the company paid dividends or not to determine the effect dividends have on this relationship.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
 ## Model set-up
+In the first model, we define $y_i$ as the stock price, $X_1$ as earnings per share, a continuous variable.
+$\beta_0$ is the intercept term, and $\beta_1$ is the coefficient for change in the stock price per dollar increase in EPS.
+$\epsilon$ denotes the noise or deviations.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+\begin{align}
+&Simple\ Model \\
+&Y_i = \beta_0 + \beta_1X_1 + \epsilon \\
+\end{align}
 
-Simple Model
+In the second model, we add $X_2$ as a binary variable where yes signifies that a company paid dividends, and $\beta_2$ is the coefficient for change in stock price if dividends were paid.
+
 \begin{align} 
-y &= \beta_0 + \beta_1X + \epsilon \\
-Y &\sim \mbox{Normal}(\beta, \sigma^2)
+&Multivariable\ Model \\
+&Y_i = \beta_0 + \beta_1X_1 + \beta_2X_2 + \epsilon \\
 \end{align}
-Where Y is the stock price and X is earnings per share.
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-Multivariable Model
-\begin{align} 
-y &= \beta_0 + \beta_1X_1 + \beta_2X_2 - \beta_3X_3 + \epsilon \\
-Y &\sim \mbox{Normal}(\beta, \sigma^2)
-\end{align}
-Where X1 is EPS, X2 is net income, and X3 is dividends.
 
 ### Model justification
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+We expect a positive linear relationship between EPS and stock price.
+This is because a higher EPS indicates greater value and investors should be willing to pay more for a company's shares if they think the company has higher profits relative to its share price.
+If a company has high EPS, they have greater amount of profit after expenses have been paid.
+Investors want to own profitable companies, so demand for this company's stock increases, driving up price as the number of stocks that are available to be bought remains the same.
+The opposite is true for unprofitable companies, if EPS is low or negative, investors may choose to sell off their shares even at a loss in order to prevent further losses should the stock price keep falling (@fig-bestfit).
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+One might expect dividends to be positively correlated with stock price because investors desire an additional payment, but the opposite is true.
+This is because dividends paid out to shareholders is value that is taken out of the company, causing price, a reflection of the firm's value, to decrease (@fig-bestfit).
+
+We expect the effects of EPS and dividends to interact with each other, creating two potential outcomes for change in price.
+Companies that did not pay dividends are expected to have a larger increase in price for every additional dollar of EPS than companies that paid dividends.
 
 # Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-first_model <-
-  lm(
-    Price ~ EPS,
-    data = analysis_data
-  )
-
-modelsummary(list("Net_Income" = first_model))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-poopy but hole
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-analysis_data <-
-  augment(
-    first_model,
-    data = analysis_data
-  )
-
-distribution_residuals <-
-  ggplot(analysis_data, aes(x = .resid)) +
-  geom_histogram(binwidth = 1) +
-  theme_classic() +
-  labs(y = "Number of occurrences", x = "Residuals")
-
-residuals_eps <-
-  ggplot(analysis_data, aes(x = EPS, y = .resid)) +
-  geom_point() +
-  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
-  theme_classic() +
-  labs(y = "Residuals", x = "Earnings Per Share")
-
-residuals_price <-
-  ggplot(analysis_data, aes(x = Price, y = .resid)) +
-  geom_point() +
-  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
-  theme_classic() +
-  labs(y = "Residuals", x = "Stock Price")
-
-estimated_vs_actual <-
-  ggplot(analysis_data, aes(x = Price, y = .fitted)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
-  theme_classic() +
-  labs(y = "Estimated Stock Price", x = "Actual Stock Price")
-
-grid.arrange(distribution_residuals, residuals_eps, residuals_price, estimated_vs_actual, nrow = 2)
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-  ggplot(analysis_data,aes(y=Price,x=EPS,color=Net_Income)) +
-  geom_point() +
-  stat_smooth(method="lm",se=FALSE)+
-  theme_classic()
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-uuhhhh
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-multi_model <-
-  lm(
-    Price ~ EPS + Net_Income,
-    data = analysis_data
-  )
-
-modelsummary(
-  list(
-    "EPS Only" = first_model,
-    "With Net Income" = multi_model
-  ),
-  fmt = 2
-)
-```
-
 ```{r}
@@
   stan_glm(
-    formula = Price ~ EPS + Net_Income,
+    formula = Price ~ EPS,
     data = analysis_data,
@@
     prior_aux = exponential(rate = 1),
-    seed = 345
+    seed = 12345
   )
@@
 #| warning: false
-
-model_rstanarm_1 <-
-  readRDS(file = "model_rstanarm.rds")
-prior_summary(model_rstanarm_1)
-```
-
-```{r}
-#| include: false
-#| eval: true
-#| warning: false
 model_rstanarm_2 <-
   stan_glm(
-    formula = Price ~ EPS + Net_Income,
+    formula = Price ~ EPS + Paid_Dividend,
     data = analysis_data,
@@
     prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 1234
+    seed = 12345
   )
@@
 ```{r}
+#| include: false
+#| eval: true
+
+summary(model_rstanarm_1) 
+summary(model_rstanarm_2)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-bestfit
+#| fig-cap: "Relationship between variables EPS and stock price"
+
+base_eps_plot <-
+  analysis_data |> 
+  ggplot(aes(x = EPS, y = Price)) +
+  geom_point(alpha = 0.5) +
+  theme_minimal() +
+  labs(x = "EPS",
+       y = "Stock Price")
+
+base_eps_plot +
+  geom_smooth(
+    method = "lm",
+    se = TRUE,
+    color = "black",
+    linetype = "dashed",
+    formula = "y ~ x"
+  )
+```
+
+In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed.
+In the first column of @tbl-modelresults, we have a intercept of -124.69 with standard deviation of 7.1 and coefficient of 21.79 with S.D.
+1.1.
+The coefficient term $\beta_1$ which shows the increase in price for every dollar increase in EPS can be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line.
+Positive slope and coefficient show a positive relationship between EPS and stock price.
+
+```{r}
 #| echo: false
 #| eval: true
-#| label: tbl-poop
-#| tbl-cap: "priors"
+#| label: fig-boxplot
+#| fig-cap: "Boxplots of stock price grouped by whether or not dividends were paid."
+#| warning: false
+
+dividends_plot <-
+  analysis_data |> 
+  ggplot(aes(x = Paid_Dividend, y = Price)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Dividends Paid",
+       y = "Stock Price")
+
+ylim1 = boxplot.stats(analysis_data$Price)$stats[c(1, 5)]
+dividends_plot = dividends_plot + coord_cartesian(ylim = ylim1*1.05)
+dividends_plot
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-model-2
+#| fig-cap: "Linear regression "
+#| warning: false
+
+base <-
+  analysis_data |>
+  ggplot(aes(x = EPS, y = Price)) +
+  labs(
+    x = "EPS",
+    y = "Price"
+  ) +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+base +
+  geom_point(aes(color = Paid_Dividend)) +
+  geom_smooth(
+    aes(color = Paid_Dividend),
+    method = "lm",
+    linetype = "dashed",
+    formula = "y ~ x"
+  ) +
+  labs(color = "Paid Dividends")
+```
+
+In model 2, we introduce dividends as an additional predictor variable.
+Stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175 (@fig-boxplot), demonstrating a negative relationship between paying dividends and stock price.
+The second column of @tbl-modelresults shows that the new intercept is 172.05 USD, and that the coefficient of paying dividends $\beta_2$ is -177.85.
+EPS of companies that paid dividends has intercept of 172.05 and coefficient of 25.59 (@tbl-modelresults).
+The new standard deviation for EPS is 1.0 and 21.6 for dividends.
+
+The effects on price from EPS and dividends can be further observed in @fig-model-2 where the observations and line of best fit are colored blue to indicate that dividends were paid and red if not.
+EPS without dividends (red line) has very similar slope and coefficient as @fig-bestfit, and EPS with dividends (blue line) is below it and significantly shorter.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Effect of earnings per share and dividends on stock price."
 #| warning: false
@@
   list(
-    "Non-scaled priors" = model_rstanarm_1,
-    "Auto-scaling priors" = model_rstanarm_2
+    "EPS Only" = model_rstanarm_1,
+    "With Dividends" = model_rstanarm_2
   ),
@@
 )
-```
-
-```{r}
-prior_summary(model_rstanarm_2)
 ```
@@
 
-# Additional data details
-
 # Model details {#sec-model-details}
@@
 In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+```{r}
+#| include: false
+#| eval: true
+
+model_rstanarm_1 <-
+  readRDS(file = "model_rstanarm.rds")
+prior_summary(model_rstanarm_1)
+```
+
+```{r}
+#| include: false
+#| eval: true
+
+prior_summary(model_rstanarm_2)
+```
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-fig-trace-1 is a trace plot
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```
+@fig-fig-trace-2 is a Rhat plot. It shows... This suggests...
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
+#| label: fig-trace
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
@@
 #| layout-ncol: 2
+
 plot(model_rstanarm_2, "trace")

### 2024-04-01 outputs/paper/paper.qmd
discussion updates

@@
 In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed.
-In the first column of @tbl-modelresults, we have a intercept of -124.69 with standard deviation of 7.1 and coefficient of 21.79 with S.D.
+In the first column of @tbl-modelresults, we have a intercept of -124.69 and coefficient of 21.79 with S.D.
 1.1.
@@
 Positive slope and coefficient show a positive relationship between EPS and stock price.
+While most observations follow the linear relationship, there is a large cluster around 0.
 
@@
 In model 2, we introduce dividends as an additional predictor variable.
-Stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175 (@fig-boxplot), demonstrating a negative relationship between paying dividends and stock price.
+Stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175 (@fig-boxplot).
 The second column of @tbl-modelresults shows that the new intercept is 172.05 USD, and that the coefficient of paying dividends $\beta_2$ is -177.85.
@@
 The effects on price from EPS and dividends can be further observed in @fig-model-2 where the observations and line of best fit are colored blue to indicate that dividends were paid and red if not.
-EPS without dividends (red line) has very similar slope and coefficient as @fig-bestfit, and EPS with dividends (blue line) is below it and significantly shorter.
+EPS without dividends (red line) has very similar slope and coefficient as @fig-bestfit, and EPS with dividends (blue line) is below.
 
@@
 # Discussion
+## Companies with Higher Earnings per Share are More Attractive to Investors
+From model 1, we obtain a coefficient of 21.79 and standard deviation of 1.1 for EPS, indicating that for every additional dollar increase in EPS, stock price increases by 21.79 USD with average deviation of 1.1 (@tbl-modelresults).
+This confirms that there is a positive linear relationship between EPS and stock price.
+When a company is profitable relative to the amount of shares they have, each shareholder is entitle to a larger portion of their profit, making the company a more attractive investment than competitors that have lower EPS.
+This demand for shares of the company then drives up stock prices.
+Alternatively, a lower EPS indicating lower profitability or losses can cause existing shareholders to sell their shares, creating downward pressure on price.
 
-## First discussion point {#sec-first-point}
+This conclusion has several implications for investors and management of companies alike.
+For investors, it may be useful to use the causal relationship between EPS and stock price to predict future stock prices.
+An investor can use available net income information and number of shares from any company to calculate EPS before the information is released to the public.
+If the estimated EPS is high, the investor can buy up shares at a lower price before the increase in demand and sell those shares once demand increases to make a profit.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Managers of a company who have an incentive for their company to be profitable in the long term would want their share prices to exhibit stable growth over time instead of the short term fluctuations caused by changing demand.
+If managers anticipates that one year's EPS will be low, the company can use excess income to buy up shares once the price decreases.
+This alleviates some of the downward pressure on price and creates more stability.
+The company can then either hold onto the purchased share until a subsequent year when the price rebounds to sell them, or eliminate those shares from the market.
+Fewer shares means that the denominator of the EPS equation is smaller, causing subsequent EPS to be higher in the future.
 
-## Second discussion point
+## Paying Dividends Causes a Decrease in Stock Price
+From @fig-boxplot there is a negative relationship between paying dividends and stock price.
+When a company chooses to pay dividends, on average their stock price is expected to decrease by about 177.85 USD with standard deviation of 21.6 USD (@tbl-modelresults).
+This is consistent with contemporary literature on dividend policy which suggests that paying dividends causes the stock price to drop by the amount of the dividend per share plus the opportunity cost of lost potential profits the dividend could have been used to generate (@CiteNarinderp).
+Dividends paid out represents value that is extracted from the company and not used to invest into profitable projects.
+All things equal, this indicates that the company is expected to grow less than competitors who did not pay dividends and reinvested, making the stock less attractive to potential investors who care about growth.
 
-## Third discussion point
+Managers are faced with the decision of what to do with extra profit.
+If paid out in dividends, the stock price will fall but existing shareholders, especially large ones with 10% or more equity in the company, would want the dividend income as compensation for their investment.
+At the same time, paying out dividends makes the stock less attractive to future investors.
+By knowing the coefficient of paying dividends, managers are better able to balance these competing priorities.
+
+EPS coefficient of companies that paid dividends is 25.59 and new standard deviation for EPS is 1.0 (@tbl-modelresults).
+This is larger than that of model 1 where dividends were not considered.
+This means that the amount price is expected to increase per dollar EPS increases is larger.
 
@@
 
-# Appendix {-}
+# Appendix
 
@@
 ## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
 ```{r}
@@
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-cap: "Examining how model 2 fits the data, and how it is affected by the data"
 #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
@@
 ```
+@fig-ppcheckandposteriorvsprior-1 compares actual dependent variable outcomes with distributions from the posterior simulation.
+We can see that the general shape and distribution are similar but actual data is far more concentrated than the simulations.
+
+@fig-ppcheckandposteriorvsprior-2 compares priors with posteriors to see how much the estimates change once data are taken into account. Our priors and posteriors are quite similar
+
+## Credible Intervals
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-credible-intervals
+#| fig-cap: "95% credibility intervals of model 2 variable coefficients"
+
+plot(
+  model_rstanarm_2,
+  "areas"
+)
+```
+
+@fig-credible-intervals shows the distributions of the 95% credibility intervals of coefficients of the predictor variables of model 2.
+The Bayesian estimation provides distributions of each coefficient.
+The actual value has a 95% probability of being within the interval with the distribution showing the likelihood within the interval.
+Ones that have a wider credibility interval are more spread out like dividends and intercept means that the true coefficient is within a larger range.
 
 ## Diagnostics
-
-@fig-fig-trace-1 is a trace plot
-
-@fig-fig-trace-2 is a Rhat plot. It shows... This suggests...
-
 ```{r}
@@
 #| warning: false
-#| label: fig-trace
+#| label: fig-trace-rhat
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
@@
 ```
+Markov chain Monte Carlo (MCMC) to was used in the model to obtain samples from the posterior distributions of interest.
+@fig-trace-rhat-1 shows bouncing horizontal lines with overlap, and @fig-trace-rhat-2 shows R-hat close to one.
+Both indicate that there are no issues with the algorithm.
 
 \newpage
-
 # References

### 2024-04-02 outputs/paper/paper.qmd
updates to discussion

@@
 ---
-title: "My title"
+title: "Relationship Between Stock Prices and Common Financial Metrics"
 subtitle: "Analysis of the Relationship Between Stock Price and EPS & Dividends of the Largest 50 North American Companies over 10 Years"
@@
 Our data comes from Compustat's North America Fundamentals Annual database, which contains both financial data from all public North American companies, collected by Compustat from each company's annual financial statements or from stock exchanges (@fig-Analysis-Data-Sample).
-From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years.
+From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years (2013-2023).
 This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr
@@
 #| warning: false
-model_rstanarm_1 <-
-  stan_glm(
-    formula = Price ~ EPS,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5),
-    prior_intercept = normal(location = 0, scale = 2.5),
-    prior_aux = exponential(rate = 1),
-    seed = 12345
+
+model_1 <-
+  lm(
+    Price ~ EPS,
+    data = analysis_data
   )
@@
 saveRDS(
-  model_rstanarm_1,
-  file = "model_rstanarm_1.rds"
+  model_1,
+  file = "model_1.rds"
 )
@@
 
-summary(model_rstanarm_1) 
+summary(model_1) 
 summary(model_rstanarm_2)
@@
 In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed.
-In the first column of @tbl-modelresults, we have a intercept of -124.69 and coefficient of 21.79 with S.D.
-1.1.
+In the first column of @tbl-modelresults, we have a intercept of 27.25 and coefficient of 26.59 with S.D.
+1.05.
 The coefficient term $\beta_1$ which shows the increase in price for every dollar increase in EPS can be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line.
@@
   list(
-    "EPS Only" = model_rstanarm_1,
+    "EPS Only" = model_1,
     "With Dividends" = model_rstanarm_2
@@
 
+EPS coefficient of companies that paid dividends is 25.59 and new standard deviation for EPS is 1.0 (@tbl-modelresults).
+This is smaller than that of model one where dividends were not considered.
+This means that the amount price is expected to increase per dollar EPS increases is larger, likely due to the fact that if a firm is profitable enough to pay dividends, it should also have high EPS.
+
 Managers are faced with the decision of what to do with extra profit.
@@
 At the same time, paying out dividends makes the stock less attractive to future investors.
-By knowing the coefficient of paying dividends, managers are better able to balance these competing priorities.
-
-EPS coefficient of companies that paid dividends is 25.59 and new standard deviation for EPS is 1.0 (@tbl-modelresults).
-This is larger than that of model 1 where dividends were not considered.
-This means that the amount price is expected to increase per dollar EPS increases is larger.
+By knowing the coefficient of paying dividends and coefficient of EPS increases, managers are better able to balance these competing priorities and make better decisions in regards to long-term profitability and growth of their firms.
 
 ## Weaknesses and next steps
+A weakness in our model is the distribution of the data where they are clustered around zero, this may indicate that a linear model is not appropriate for our observations and the conclusions between EPS, dividends and price cannot be drawn.
+While the residuals follow normal distribution, they are clustered together instead of randomly spaced apart, indicating that they contain some additional structure that was not captured by our analysis (@fig-residuals).
+The posterior prediction check also supports this as the actual data is far larger kurtosis than what the model predicted, meaning that the predictive power of the model is low (@fig-ppcheckandposteriorvsprior-1).
+The credibility intervals for the coefficient of paying dividends is also quite large, with probability normally distributed between about -275 and -125 USD, suggesting a relatively high level of uncertainty (@fig-credible-intervals).
+Practically, the model worked as intended with no signs of problems from the trace plot or r-hat values (@fig-trace-rhat).
+Further work is needed to confirm the positive linear relationship between stock price and EPS.
 
-Weaknesses and next steps should also be included.
+The conclusions drawn from our analysis only applies to our data set of large, North American companies from 2013-2023.
+More analysis with diverse data sets in terms of need to be done to determine if the same conclusions are applicable to companies of different sizes or different geographic locations.
+The data is also historic, meaning that the relationships described may not occur in the future.
+Training the model to have predictive power may provide additional useful information.
+
+Lastly, while financial metrics are important to management and investor decision making, there are many quantitative factors like the strategic direction of a firm and its competitive advantage within the industry should also be considered.
+Future analysis should aim to quantify these categorical factors and determine their relationship with stock price as well as how they interact with financial metrics.
 
@@
 
-## Posterior predictive check
-```{r}
-#| include: false
-#| eval: true
-
-model_rstanarm_1 <-
-  readRDS(file = "model_rstanarm.rds")
-prior_summary(model_rstanarm_1)
-```
-
 ```{r}
@@
 
+## Residuals
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-residuals
+#| layout-ncol: 2
+#| layout-nrow: 2
+#| fig-cap: "Examining how model 2 fits the data, and how it is affected by the data"
+#| fig-subcap: ["Distribution of residuals", " Residuals by EPS", "Residuals by Price", "Comparing the estimated price with the actual price"]
+
+analysis_data <-
+  augment(
+    model_1,
+    data = analysis_data
+  )
+
+ggplot(analysis_data, aes(x = .resid)) +
+  geom_histogram(binwidth = 1) +
+  theme_classic() +
+  labs(y = "Number of occurrences", x = "Residuals")
+
+ggplot(analysis_data, aes(x = EPS, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
+  theme_classic() +
+  labs(y = "Residuals", x = "EPS")
+
+ggplot(analysis_data, aes(x = Price, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
+  theme_classic() +
+  labs(y = "Residuals", x = "Stock Price")
+
+ggplot(analysis_data, aes(x = Price, y = .fitted)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_classic() +
+  labs(y = "Estimated Stock Price", x = "Actual Stock Price")
+```
+
+## Posterior predictive check
 ```{r}
@@
 @fig-ppcheckandposteriorvsprior-1 compares actual dependent variable outcomes with distributions from the posterior simulation.
-We can see that the general shape and distribution are similar but actual data is far more concentrated than the simulations.
+We can see that the general shape and distribution are similar but actual data has far higher kurtosis than the simulations.
 
-@fig-ppcheckandposteriorvsprior-2 compares priors with posteriors to see how much the estimates change once data are taken into account. Our priors and posteriors are quite similar
+@fig-ppcheckandposteriorvsprior-2 compares priors with posteriors to see how much the estimates change once data are taken into account. Our priors and posteriors are quite similar,
 

### 2024-04-02 outputs/paper/paper.qmd
title
link

@@
 author: Diana Liu
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Relationship-Between-Stock-Prices-and-Common-Financial-Metrics"
 date: today
@@
 # Introduction
-
-You can and should cross-reference sections and sub-sections.
-
-The remainder of this paper is structured as follows. @sec-data....
 
@@
   model_1,
-  file = "model_1.rds"
+  file = "~/Stock-Prices/outputs/models/model_1.rds"
 )
@@
   model_rstanarm_2,
-  file = "model_rstanarm_2.rds"
+  file = "~/Stock-Prices/outputs/models/model_rstanarm_2.rds"
 )

### 2024-04-06 outputs/paper/paper.qmd
update sketches

@@
 library(knitr)
-#
+library(grDevices)
 library(gt)
-library(grDevices)
 ```
@@
 From this data set, we extract the price, EPS, net income, dividends, TIC, and Year of the top 50 largest companies by market capitalization, a measure of the size of the company in terms of total value, for the past 10 years (2013-2023).
-This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr
+This Data is cleaned and analysed in @CiteR with assistance from @CiteTidy, @CiteArrow, @CiteModelsummary, @CiteBayesplot, @CiteBroom, @CiteGridextra, @CiteRstanarm, @CiteKnitr, @Citegt, and @CitegrDevices.
 

### 2024-04-07 outputs/paper/paper.qmd
paper updates

@@
     formula = "y ~ x"
-  )
+  ) +
+  xlim(-5, 30) +
+  ylim(0, 2000)
 ```
@@
 In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed.
-In the first column of @tbl-modelresults, we have a intercept of 27.25 and coefficient of 26.59 with S.D.
-1.05.
-The coefficient term $\beta_1$ which shows the increase in price for every dollar increase in EPS can be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line.
+In the first column of @tbl-modelresults, we have a intercept of 27.25 and with standard error 11.47.
+The coefficient term $\beta_1$ is 26.59 with standard error 1.05 which shows the increase in price for every dollar increase in EPS that can also be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line.
 Positive slope and coefficient show a positive relationship between EPS and stock price.
-While most observations follow the linear relationship, there is a large cluster around 0.
+While most observations follow the linear relationship, there is some clustering between 0 to 10.
 
@@
 #| label: fig-model-2
-#| fig-cap: "Linear regression "
+#| fig-cap: "Relationship between EPS, dividends, and stock price."
 #| warning: false
@@
   ) +
-  labs(color = "Paid Dividends")
+  labs(color = "Paid Dividends") +
+  xlim(-5, 30) +
+  ylim(0, 2000)
 ```
@@
 In model 2, we introduce dividends as an additional predictor variable.
-Stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175 (@fig-boxplot).
+Paying dividends has a negative relationship with price as show in @fig-boxplot where stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175.
+Stocks that did not pay dividends also have a significantly larger interquartile range for price of around 250 USD than stocks that did of 100.
+
 The second column of @tbl-modelresults shows that the new intercept is 172.05 USD, and that the coefficient of paying dividends $\beta_2$ is -177.85.
-EPS of companies that paid dividends has intercept of 172.05 and coefficient of 25.59 (@tbl-modelresults).
-The new standard deviation for EPS is 1.0 and 21.6 for dividends.
-
+With the addition of the dividends variable, EPS of companies that paid dividends has changed to a coefficient of 25.59 (@tbl-modelresults).
 The effects on price from EPS and dividends can be further observed in @fig-model-2 where the observations and line of best fit are colored blue to indicate that dividends were paid and red if not.
-EPS without dividends (red line) has very similar slope and coefficient as @fig-bestfit, and EPS with dividends (blue line) is below.
+EPS without dividends (red line) has very similar slope and coefficient as @fig-bestfit.
+EPS with dividends (blue line) has a less steep slope and is located below the EPS without dividends.
 
@@
 ```
+$R^2$ and adjusted $R^2$ are terms that explains how well the model is able to explain data, with $R^2$ for single variable models and the more conservative adjusted $R^2$ for multiple variables.
+Model 1 has a $R^2$ of 0.57, meaning that 57% of the variability observed in stock price is explained by EPS in the regression model.
+Model 2 can explain 59% of stock price with EPS and dividends.
+Both indicate that the model can explain stock price reasonably well, but there is around 40% of variability is attributed to other causes.
 
@@
 ## Companies with Higher Earnings per Share are More Attractive to Investors
-From model 1, we obtain a coefficient of 21.79 and standard deviation of 1.1 for EPS, indicating that for every additional dollar increase in EPS, stock price increases by 21.79 USD with average deviation of 1.1 (@tbl-modelresults).
+From model 1, we obtain a coefficient of 26.59 and standard error of 1.05 for EPS, indicating that for every additional dollar increase in EPS, stock price increases by 26.599 USD (@tbl-modelresults).
+The standard error term means that on average the relationship between EPS and stock price differs from the actual relationship by 1.05 USD.
+The standard error is relatively small, just over a dollar, so model 1 is reasonably accurate.
 This confirms that there is a positive linear relationship between EPS and stock price.
@@
 
-## Paying Dividends Causes a Decrease in Stock Price
+## Paying Dividends Causes Stock Prices to Fall
 From @fig-boxplot there is a negative relationship between paying dividends and stock price.
-When a company chooses to pay dividends, on average their stock price is expected to decrease by about 177.85 USD with standard deviation of 21.6 USD (@tbl-modelresults).
+When a company chooses to pay dividends, on average their stock price is expected to decrease by about 177.85 USD (@tbl-modelresults).
 This is consistent with contemporary literature on dividend policy which suggests that paying dividends causes the stock price to drop by the amount of the dividend per share plus the opportunity cost of lost potential profits the dividend could have been used to generate (@CiteNarinderp).
@@
 
-EPS coefficient of companies that paid dividends is 25.59 and new standard deviation for EPS is 1.0 (@tbl-modelresults).
-This is smaller than that of model one where dividends were not considered.
-This means that the amount price is expected to increase per dollar EPS increases is larger, likely due to the fact that if a firm is profitable enough to pay dividends, it should also have high EPS.
+EPS coefficient of companies that paid dividends is 25.59 which is one dollar less than that of model one where dividends were not considered (@tbl-modelresults).
+This means that the amount price is expected to increase per dollar EPS increases is smaller, likely due to  !!
+why lol?<<<
 
@@
 #| layout-nrow: 2
-#| fig-cap: "Examining how model 2 fits the data, and how it is affected by the data"
+#| fig-cap: "Residuals of model 1, simple linear regression"
 #| fig-subcap: ["Distribution of residuals", " Residuals by EPS", "Residuals by Price", "Comparing the estimated price with the actual price"]
@@
   theme_classic() +
-  labs(y = "Residuals", x = "EPS")
+  labs(y = "Residuals", x = "EPS") +
+  xlim(-5, 30) +
+  ylim(-500, 2000)
 
@@
   theme_classic() +
-  labs(y = "Residuals", x = "Stock Price")
+  labs(y = "Residuals", x = "Stock Price") +
+  xlim(0, 2000) +
+  ylim(-500, 2000)
 
@@
   theme_classic() +
-  labs(y = "Estimated Stock Price", x = "Actual Stock Price")
+  labs(y = "Estimated Stock Price", x = "Actual Stock Price") +
+  xlim(0, 2000) +
+  ylim(-500, 2000)
 ```
+@fig-residuals-1 shows that residuals are normally distributed around zero which is a good sign that our model is able to capture the main patterns in stock price data with our predictor variables.
+This is supported by residuals scatter plot @fig-residuals-3 show that the residual values are randomly scattered around zero.
+However, in @fig-residuals-2, the residuals show a negative linear pattern which indicates that our linear model may not be appropriate for the data.
+Lastly, in @fig-residuals-4, the relationship between actual and estimated stock prices appears somewhat linear but with a large cluster around 0.
+A perfectly predictive model will have a linear slope of one, meaning that our model may lack predictive power.
 
@@
 #| label: fig-trace-rhat
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-cap: "Checking the convergence of the MCMC algorithm of model 2"
 #| fig-subcap: ["Trace plot", "Rhat plot"]

### 2024-04-12 outputs/paper/paper.qmd
paper updates

@@
 ---
-title: "Relationship Between Stock Prices and Common Financial Metrics"
-subtitle: "Analysis of the Relationship Between Stock Price and EPS & Dividends of the Largest 50 North American Companies over 10 Years"
+title: "Poopy But hole"
+subtitle: "Analysis of the Relationship Between Stock Price, Earnings Per Share, and Dividends of the Largest 50 North American Companies from 2013 to 2023"
 author: Diana Liu
@@
 Both indicate that the model can explain stock price reasonably well, but there is around 40% of variability is attributed to other causes.
+This may also be why the @fig-model-2 does not look perfectly linear and has clustering around 0 to 10 EPS.
 
@@
 ## Companies with Higher Earnings per Share are More Attractive to Investors
-From model 1, we obtain a coefficient of 26.59 and standard error of 1.05 for EPS, indicating that for every additional dollar increase in EPS, stock price increases by 26.599 USD (@tbl-modelresults).
-The standard error term means that on average the relationship between EPS and stock price differs from the actual relationship by 1.05 USD.
-The standard error is relatively small, just over a dollar, so model 1 is reasonably accurate.
+From model 1, we obtain an intercept of 27.25 meaning that when EPS is zero, stock price is 27.35 USD.
+This term has standard error of 11.47USD which means that the estimate is not very precise as it usually deviates from the real stock price by 11.47.
+We also obtain a coefficient of 26.59 with standard error of 1.05 for EPS, indicating that for every additional dollar increase in EPS, stock price increases by 26.599 USD (@tbl-modelresults).
+The standard error is relatively small, just over a dollar, so model 1 is reasonably accurate for predicting stock price using.
 This confirms that there is a positive linear relationship between EPS and stock price.
@@
 When a company chooses to pay dividends, on average their stock price is expected to decrease by about 177.85 USD (@tbl-modelresults).
-This is consistent with contemporary literature on dividend policy which suggests that paying dividends causes the stock price to drop by the amount of the dividend per share plus the opportunity cost of lost potential profits the dividend could have been used to generate (@CiteNarinderp).
+This is consistent with contemporary literature on dividend policy which suggests that paying dividends causes the stock price to drop by the amount of the dividend per share plus the opportunity cost of lost potential profits the dividend could have been used to generate [@CiteNarinderp].
 Dividends paid out represents value that is extracted from the company and not used to invest into profitable projects.
-All things equal, this indicates that the company is expected to grow less than competitors who did not pay dividends and reinvested, making the stock less attractive to potential investors who care about growth.
+All things equal, this indicates that the company is expected to grow less than competitors who did not pay dividends and reinvested, making the stock less attractive to investors who care about growth.
 
 EPS coefficient of companies that paid dividends is 25.59 which is one dollar less than that of model one where dividends were not considered (@tbl-modelresults).
-This means that the amount price is expected to increase per dollar EPS increases is smaller, likely due to  !!
-why lol?<<<
+This means that the amount price is expected to increase per dollar EPS increases is smaller, likely due to the negative stock price effect of dividends interacting with the positive effect of EPS.
+Since the EPS coefficient is only reduced by a dollar, we can assume that the positive EPS effect is stronger than that of the dividends.
 
-Managers are faced with the decision of what to do with extra profit.
-If paid out in dividends, the stock price will fall but existing shareholders, especially large ones with 10% or more equity in the company, would want the dividend income as compensation for their investment.
+These results can be useful for managers who are faced with the decision of what to do with extra profit.
+If paid out in dividends, the stock price will fall, but existing shareholders, especially large ones with 10% or more equity in the company, would want the dividend income as compensation for their investment and push management to pay dividends.
 At the same time, paying out dividends makes the stock less attractive to future investors.
-By knowing the coefficient of paying dividends and coefficient of EPS increases, managers are better able to balance these competing priorities and make better decisions in regards to long-term profitability and growth of their firms.
+By knowing the coefficient of paying dividends and how it effects the EPS coefficient, managers are better able to balance these competing priorities and make better decisions in regards to long-term profitability and growth of their firms.
+
+Investors can use these results to secure profit opportunities.
+If a company announces that they will pay dividends, investors can anticipate the decline in stock price and react accordingly, either buying the stock to receive the dividend or selling and locking in returns before the price drops.
+How paying dividends decrease the positive effect on price EPS has can also be used to their advantage.
+Based on EPS and dividends information which are publicly available, investors can more accurately predict the future stock price of any given company and make a profit by buying or selling its stocks.
 
@@
 Practically, the model worked as intended with no signs of problems from the trace plot or r-hat values (@fig-trace-rhat).
-Further work is needed to confirm the positive linear relationship between stock price and EPS.
+Further work is needed to confirm the positive linear relationship between stock price and EPS, as well as how paying dividends affect this relationship.
 
 The conclusions drawn from our analysis only applies to our data set of large, North American companies from 2013-2023.
-More analysis with diverse data sets in terms of need to be done to determine if the same conclusions are applicable to companies of different sizes or different geographic locations.
-The data is also historic, meaning that the relationships described may not occur in the future.
-Training the model to have predictive power may provide additional useful information.
+More analysis with diverse data sets need to be done to determine if the same conclusions are applicable to companies of different sizes or different geographic locations.
+The data is also historic, meaning that the relationships described may not hold in the future.
+Extraneous events like the pandemic are likely to effect the relationships studied in ways that cannot be predicted by our model [@CiteBasunoy].
 
-Lastly, while financial metrics are important to management and investor decision making, there are many quantitative factors like the strategic direction of a firm and its competitive advantage within the industry should also be considered.
-Future analysis should aim to quantify these categorical factors and determine their relationship with stock price as well as how they interact with financial metrics.
+Lastly, while financial metrics are important to management and investor decision making, there are many qualitative factors that should also be considered like the strategic direction of a firm, the strength of its customer relations and workforce productivity, and its competitive advantage within the industry.
+The WRDS database we used did not have metrics to measure these variables in terms of numerical value, so ways to quantify these variables should be explored.
+Future analysis should aim to determine the relationship between these categorical variables with stock price as well as how they interact with other financial metrics like EPS and dividends.
+
+# Conclusion
+We have determined that there is a positive linear relationship between EPS and stock price, and a negative relationship between dividends and stock price which reduces the affect of EPS on price.
+These relationships can be used by investors to predict stock price of a given company based on EPS and dividends.
+The investor can then earn money by trading stocks accordingly, buying when the price is predicted to increase and selling when it will fall.
+Managers of companies can also use our results to make more informed decisions, balancing the objectives of pleasing current shareholders, investing excess cash into promising projects, and maintainging the attractiveness of their stock for new investors.
 
 \newpage
-
 \appendix
-
 # Appendix
-
 # Model details {#sec-model-details}

### 2024-04-12 scripts/00-simulate_data.R

@@ -15,17 +15,15 @@ tic1 <- sample(letters, 100, TRUE)
 tic2 <- sample(letters, 100, TRUE)
 tic3 <- sample(letters, 100, TRUE)
 tic <- paste(tic1, tic2, tic3)
+dividend <- c("yes", "no")
 
 financial_data <-
   tibble(
     tic,
     net_income = round(abs(rnorm(100, mean = 10000, sd = 5000)), 2),
-    dividends = round(abs(rnorm(100, mean = 5, sd = 100)), 2),
+    dividends = sample(dividend, 100, replace = T, prob = c(0.5, 0.5)),
     eps = round(abs(rnorm(100, mean = 5, sd = 2)), 2),
     price = round(abs(rnorm(100, mean = 75, sd = 100)), 2)
   )
 
 financial_data
-
-
-

### 2024-04-12 outputs/paper/paper.qmd
intro and abstract

@@
 ---
-title: "Poopy But hole"
+title: "Effect of Earnings and Dividends on Stock Prices"
 subtitle: "Analysis of the Relationship Between Stock Price, Earnings Per Share, and Dividends of the Largest 50 North American Companies from 2013 to 2023"
@@
 # Introduction
+Public companies whose stock are traded on exchanges, must disclose certain financial metrics in their quarterly and yearly financial statements.
+Investors can then use these metrics to help inform their decision when selecting companies to invest in, their investment of buying shares represent a fractional ownership of the company.
+Some of the most important metrics investors look at are Net Income, Earnings Per Share (EPS), and Dividends [@CiteShakespeare].
+In the data section, we discuss how EPS is a general measure of a firm's profitability, dividends are cash distributed to investors, and stock prices are determined by both the intrinsic value of the firm as well as market pressures like demand for the stock.
+Our estimand is the relationship between EPS and stock price as well has how this relationship is effected by dividends.
+
+In the model section, we construct two models with model 1 being a simple linear model between EPS and stock price to determine their relationship, and model 2 being a multivariable model that groups companies by those who paid dividends and those who did not.
+From the results section, model 1 found that there is a positive linear relationship between EPS and stock price with every dollar increase in EPS correlating with a 26.59 USD increase in stock price.
+We also determined that there is a negative relationship between paying dividends and stock price.
+In Model 2, if the company paid dividends, every dollar increase in EPS correlates with only a 25.59 USD increase in stock price.
+
+We have determined that there is a positive linear relationship between EPS and stock price, and a negative relationship between dividends and stock price which reduces the affect of EPS on price.
+These relationships can be used by investors to predict stock price of a given company based on EPS and dividends.
+The investor can then earn money by trading stocks accordingly, buying when the price is predicted to increase and selling when it will fall.
+Managers of companies can also use our results to make more informed decisions, balancing the objectives of pleasing current shareholders, investing excess income into promising projects, and maintaining the attractiveness of their stock for new investors.
+Further analysis should be conducted using a more diverse data set in terms of geography and company size in order to confirm these relationships for all companies.
+Another weakness is that stock price is also influenced by categorical factors like a company's strategic direction and competitive advantage which are variables that could not be included in our analysis.
+Future efforts should attempt to determine the effects of these categorical variables on stock price and how it interacts with that of EPS and dividends.
 
 # Data
-Public companies whose stock are traded on exchanges, must disclose certain financial metrics in their quarterly and yearly financial statements.
-Investors then use these metrics to help inform their decision when selecting companies to invest in.
-Some of the most important metrics investors look at are Earnings Per Share, Net Income, and Dividends [@CiteShakespeare].
-Investments take the form of buying shares or stocks which allows the investor to own a fraction of the company.
-
 The data used for analysis is obtained through Walton Research Data Services (WRDS) by the University of Pennsylvania.
@@
 # Conclusion
-We have determined that there is a positive linear relationship between EPS and stock price, and a negative relationship between dividends and stock price which reduces the affect of EPS on price.
-These relationships can be used by investors to predict stock price of a given company based on EPS and dividends.
-The investor can then earn money by trading stocks accordingly, buying when the price is predicted to increase and selling when it will fall.
-Managers of companies can also use our results to make more informed decisions, balancing the objectives of pleasing current shareholders, investing excess cash into promising projects, and maintainging the attractiveness of their stock for new investors.
 

### 2024-04-12 outputs/paper/paper.qmd
abstract

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study examines the relationship between Earnings per Share (EPS), dividends, and stock prices in publicly traded companies.
+Findings indicate a positive correlation between EPS and stock price, with each dollar increase in EPS associated with a $26.59 rise in stock price.
+Conversely, dividend payments are linked to a decrease in stock price, mitigating the impact of EPS on price.
+These insights enable investors to predict stock prices, informing their investment strategies, as well as guide managerial decisions balancing pleasing existing shareholders by paying dividends, ensuring long-term growth of the company, and keeping the stock attractive to new investors."
 format: pdf
@@
 # Conclusion
+The analysis indicates a positive linear relationship between Earnings per Share (EPS) and stock price, suggesting that companies with higher EPS are more attractive to investors.
+This relationship allows investors to predict future stock prices based on EPS data before it's publicly released, providing opportunities for profit.
+However, paying dividends is associated with a decrease in stock price, likely due to the perceived decrease in growth potential when profits are distributed rather than reinvested.
+Managers facing the decision of dividend payout must balance the desires of existing shareholders for income against the potential negative impact on future investor attraction.
+The model's weaknesses, such as clustered data distribution and high uncertainty in dividend coefficient estimates, suggest the need for further analysis.
+Exploration of qualitative factors beyond financial metrics to better understand stock price dynamics can also be beneficial to investors and managers.
+Overall, while EPS remains a significant predictor of stock price, future studies should aim to incorporate a broader range of variables and data sets to enhance predictive accuracy and account for potential variations across industries and time periods.
 

### 2024-04-12 outputs/paper/paper.qmd
paper updates

@@
 author: Diana Liu
-thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Relationship-Between-Stock-Prices-and-Common-Financial-Metrics"
+thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Effect-of-Earnings-and-Dividends-on-StockPrices"
 date: today

### 2024-04-13 outputs/paper/paper.qmd
update paper and modeling script

@@
 ---
-title: "Effect of Earnings and Dividends on Stock Prices"
+title: "What Determines Stock Prices?"
 subtitle: "Analysis of the Relationship Between Stock Price, Earnings Per Share, and Dividends of the Largest 50 North American Companies from 2013 to 2023"
 author: Diana Liu
-thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Effect-of-Earnings-and-Dividends-on-StockPrices"
+thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/What-Determines-Stock-Prices-"
 date: today
@@
 
-model_1 <-
-  lm(
-    Price ~ EPS,
-    data = analysis_data
-  )
-
-saveRDS(
-  model_1,
-  file = "~/Stock-Prices/outputs/models/model_1.rds"
-)
-```
-
-```{r}
-#| include: false
-#| eval: true
-#| warning: false
-model_rstanarm_2 <-
-  stan_glm(
-    formula = Price ~ EPS + Paid_Dividend,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(0, 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 12345
-  )
-
-saveRDS(
-  model_rstanarm_2,
-  file = "~/Stock-Prices/outputs/models/model_rstanarm_2.rds"
-)
+model_1 <- readRDS("~/Stock-Prices/outputs/models/model_1.rds")
+model_2 <- readRDS("~/Stock-Prices/outputs/models/model_2.rds")
 ```
@@
 summary(model_1) 
-summary(model_rstanarm_2)
+summary(model_2)
 ```
@@
     "EPS Only" = model_1,
-    "With Dividends" = model_rstanarm_2
+    "With Dividends" = model_2
   ),
@@
 
-prior_summary(model_rstanarm_2)
+prior_summary(model_2)
 ```
@@
 
-pp_check(model_rstanarm_2) +
+pp_check(model_2) +
   theme_classic() +
@@
 
-posterior_vs_prior(model_rstanarm_2) +
+posterior_vs_prior(model_2) +
   theme_minimal() +
@@
 plot(
-  model_rstanarm_2,
+  model_2,
   "areas"
@@
 
-plot(model_rstanarm_2, "trace")
+plot(model_2, "trace")
 
-plot(model_rstanarm_2, "rhat")
+plot(model_2, "rhat")
 ```

### 2024-04-13 scripts/03-test_data.R

@@ -13,7 +13,7 @@ library(tidyverse)
 #### Test data ####
 analysis_data <- read_parquet("outputs/data/analysis_data.parquet")
 
-# Tests if the ticker symbol is a character
+# 1. Tests if the ticker symbol is a character
 if (all(analysis_data$Tic |>
         unique() %>% class() == "character")) {
   "Pass: The cleaned Tic symbols are the correct class"
@@ -21,7 +21,7 @@ if (all(analysis_data$Tic |>
   "Fail: Not all of the Tic symbols have been cleaned completely"
 }
 
-# Tests if the ticker symbol has 6 or less characters
+# 2. Tests if the ticker symbol has 6 or less characters
 if (all(analysis_data$Tic |>
         nchar() <= 6)) {
   "Pass: The cleaned Tic symbols are the correct length"
@@ -29,7 +29,7 @@ if (all(analysis_data$Tic |>
   "Fail: Not all of the Tic symbols have been cleaned completely"
 }
 
-# Tests if the Year is positive
+# 3. Tests if the Year is positive
 if (all(analysis_data$Year |>
         min() > 0)) {
   "Pass: The cleaned Years are positive"
@@ -37,7 +37,7 @@ if (all(analysis_data$Year |>
   "Fail: Years have not been cleaned completely"
 }
 
-# Tests if the Year is greater than or equal to 2013
+# 4. Tests if the Year is greater than or equal to 2013
 if (all(analysis_data$Year |>
         min() >= 2013)) {
   "Pass: The cleaned Years are greater than or equal to 2013"
@@ -45,7 +45,7 @@ if (all(analysis_data$Year |>
   "Fail: Years have not been cleaned completely"
 }
 
-# Tests if the Year is less than or equal to 2023
+# 5. Tests if the Year is less than or equal to 2023
 if (all(analysis_data$Year |>
         max() <= 2023)) {
   "Pass: The cleaned Years are less than or equal to 2023"
@@ -53,7 +53,7 @@ if (all(analysis_data$Year |>
   "Fail: Years have not been cleaned completely"
 }
 
-# Tests if Dividends are numbers
+# 6. Tests if Dividends are numbers
 if (all(analysis_data$Dividends |>
         unique() %>% class() == "numeric")) {
   "Pass: The cleaned Dividends are the correct class"
@@ -61,7 +61,7 @@ if (all(analysis_data$Dividends |>
[…]

### 2024-04-13 scripts/04-model.R

@@ -1,37 +1,47 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Models stock price data with a linear model
+# Author: Diana Liu
+# Date: 11 April 12 2024
+# Contact: guanzhi.liu@mail.utoronto.ca
+# Pre-requisites: downloaded, cleaned, and tested data
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("outputs/data/analysis_data.csv")
+analysis_data <- read_parquet(here::here("outputs/data/analysis_data.parquet"))
 
 ### Model data ####
-first_model <-
+model_1 <-
+  lm(
+    Price ~ EPS,
+    data = analysis_data
+  )
+
+model_2 <-
   stan_glm(
-    formula = flying_time ~ length + width,
+    formula = Price ~ EPS + Paid_Dividend,
     data = analysis_data,
     family = gaussian(),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(0, 2.5, autoscale = TRUE),
     prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+    seed = 12345
   )
-
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "outputs/models/first_model.rds"
+  model_1,
+  file = "~/Stock-Prices/outputs/models/model_1.rds"
 )
+
+saveRDS(
+  model_2,
+  file = "~/Stock-Prices/outputs/models/model_2.rds"
+)
+
 
 

### 2024-04-18 scripts/00-simulate_data.R
update script annotations

@@ -11,12 +11,16 @@ library(dplyr)
 library(knitr)
 
 #### Simulate data ####
+#simulate the TIC symbol
 tic1 <- sample(letters, 100, TRUE)
 tic2 <- sample(letters, 100, TRUE)
 tic3 <- sample(letters, 100, TRUE)
 tic <- paste(tic1, tic2, tic3)
+
+#simulate binary dividend variable
 dividend <- c("yes", "no")
 
+#simulate rest of data
 financial_data <-
   tibble(
     tic,

### 2024-04-18 scripts/02-data_cleaning.R

@@ -9,7 +9,6 @@
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
-#install.packages("arrow")
 library(arrow)
 
 #### Clean data ####

### 2024-04-18 scripts/03-test_data.R

@@ -8,7 +8,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
-
+library(arrow)
 
 #### Test data ####
 analysis_data <- read_parquet("outputs/data/analysis_data.parquet")

### 2024-04-18 scripts/04-model.R

@@ -15,12 +15,16 @@ library(arrow)
 analysis_data <- read_parquet(here::here("outputs/data/analysis_data.parquet"))
 
 ### Model data ####
+#create simple linear regression model with only EPS as predictor variable
+#and stock price as dependent variable
 model_1 <-
   lm(
     Price ~ EPS,
     data = analysis_data
   )
 
+#create model with EPS and binary dividend variable as predictor variables and
+#Stock price as dependent variable
 model_2 <-
   stan_glm(
     formula = Price ~ EPS + Paid_Dividend,

### 2024-04-18 outputs/paper/paper.qmd
updates to figure captions

@@
 ---
-title: "What Determines Stock Prices?"
+title: "Earnings per Share and Dividends Determine Stock Prices"
 subtitle: "Analysis of the Relationship Between Stock Price, Earnings Per Share, and Dividends of the Largest 50 North American Companies from 2013 to 2023"
 author: Diana Liu
-thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/What-Determines-Stock-Prices-"
+thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Earnings-per-Share-and-Dividends-Determine-Stock-Prices"
 date: today
@@
 Investors can then use these metrics to help inform their decision when selecting companies to invest in, their investment of buying shares represent a fractional ownership of the company.
-Some of the most important metrics investors look at are Net Income, Earnings Per Share (EPS), and Dividends [@CiteShakespeare].
+Some of the most important metrics investors look at are Net Income, Earnings Per Share (EPS), and Dividends, but whether or not these metrics are actually able to be used to predict stock prices have been disputed [@CiteShakespeare].
 In the data section, we discuss how EPS is a general measure of a firm's profitability, dividends are cash distributed to investors, and stock prices are determined by both the intrinsic value of the firm as well as market pressures like demand for the stock.
@@
 
-```{r}
+```{r, out.width="75%"}
 #| label: fig-EPS-boxplot
-#| fig-cap: "Boxplots of EPS grouped by dividends"
+#| fig-cap: "Boxplots of EPS grouped by dividends.
+EPS of companies that did not pay dividends has a larger interquartile range than EPS of companies that paid dividends.
+The mean and median EPS of companies that paid dividends is higher, suggesting that paying dividends is correlated with having higher EPS."
 #| echo: false
@@
 EPS_boxplot
-
 ```
@@
 #| label: fig-distributions
-#| fig-cap: "Price, EPS, Net_Income, and Dividends are all normally distributed with right skew"
+#| fig-cap: "Price, EPS, Net_Income, and Dividends are all normally distributed with right skew."
 #| echo: false
@@
 Investors want to own profitable companies, so demand for this company's stock increases, driving up price as the number of stocks that are available to be bought remains the same.
-The opposite is true for unprofitable companies, if EPS is low or negative, investors may choose to sell off their shares even at a loss in order to prevent further losses should the stock price keep falling (@fig-bestfit).
+The opposite is true for unprofitable companies, if EPS is low or negative, investors may choose to sell off their shares even at a loss in order to prevent further losses should the stock price keep falling.
 
 One might expect dividends to be positively correlated with stock price because investors desire an additional payment, but the opposite is true.
-This is because dividends paid out to shareholders is value that is taken out of the company, causing price, a reflection of the firm's value, to decrease (@fig-bestfit).
+This is because dividends paid out to shareholders is value that is taken out of the company, causing price, a reflection of the firm's value, to decrease.
 
@@
 
+In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed.
+In the first column of @tbl-modelresults, we have a intercept of 27.25 and with standard error 11.47.
+The coefficient term $\beta_1$ is 26.59 with standard error 1.05 which shows the increase in price for every dollar increase in EPS.
+The slope can also be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line.
+Positive slope and coefficient show a positive relationship between EPS and stock price.
+While most observations follow the linear relationship, there is some clustering between 0 to 10.
+
 ```{r}
@@
 #| label: fig-bestfit
-#| fig-cap: "Relationship between variables EPS and stock price"
+#| fig-cap: "Relationship between variables EPS and stock price.
+The coefficient or slope can be observed as the dashed line with standard error as the shaded area around the dashed line.
+Positive slope and coefficient show a positive linear relationship between EPS and stock price.
+While most observations follow the linear relationship, there is some clustering between 0 to 10."
 
@@
   geom_point(alpha = 0.5) +
-  theme_minimal() +
+  theme_classic() +
   labs(x = "EPS",
@@
 
-In model 1, only the relationship between the dependent variable price and predictor variable EPS is analyzed.
-In the first column of @tbl-modelresults, we have a intercept of 27.25 and with standard error 11.47.
-The coefficient term $\beta_1$ is 26.59 with standard error 1.05 which shows the increase in price for every dollar increase in EPS that can also be observed from @fig-bestfit as the dashed line with standard error as the shaded area around the dashed line.
-Positive slope and coefficient show a positive relationship between EPS and stock price.
-While most observations follow the linear relationship, there is some clustering between 0 to 10.
-
-```{r}
+```{r, out.width="75%"}
 #| echo: false
@@
 #| label: fig-model-2
-#| fig-cap: "Relationship between EPS, dividends, and stock price."
+#| fig-cap: "Relationship between EPS, dividends, and stock price.
+Paying dividends has a negative relationship with price as stocks that paid dividends have median price around 100, significantly lower than stocks that paid no dividends with median price around 175.
+Stocks that did not pay dividends also have a significantly larger interquartile range for price of around 250 USD than stocks that did of 100."
 #| warning: false
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Effect of earnings per share and dividends on stock price."
+#| tbl-cap: "Effect of earnings per share and dividends on stock price.
+In model 1, there is an intercept of 27.25 and with standard error 11.47.
+The coefficient term is 26.59 with standard error 1.05.
+In model 2, the new intercept is 172.05 USD, and that the coefficient of paying dividends is -177.85.
+With the addition of the dividends variable, EPS of companies that paid dividends has changed to a coefficient of 25.59 "
 #| warning: false
@@
 ## Credible Intervals
-```{r}
+```{r, out.width="75%"}
 #| eval: true
